---
title: "MAG_plotting"
author: "Ioannis D. Kampouris"
date: "2025-03-05"
output: pdf_document
---

Please do not forget to change the path to the Dereplicated Genomes.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Run PROKKA in parallel**
ls *.fasta | parallel --verbose "prokka {} --prefix {.}_out --cpus 1"

**Run checkm2** 
checkm2  predict -i dereplicated_genomes -o  checked_genomes --threads 10 -x fasta

**classification was run with the code from GTDBtk Release 2020 in Conda**
gtdbtk classify_wf input_folder output_folder

***Run antismash***
ls *fasta | parallel --verbose "antismash {} --output-dir antismash2{.} --fullhmmer --clusterhmmer --tigrfam --cc-mibig  --cpus 1 --genefinding-tool prodigal"

**Diamond anotations with PLABASE**
ls *out/*faa | parallel --verbose "diamond blastp --query {}  --out {}pl_diammond_.txt --db /home/ioannis.kampouris/three_years_CTEXT/diamond_the/plantdb.dmnd --id 90 --query-cover 50  -k 1 -e 0.000000001  -p 2"

***Now in R***
Load the MAGs
First the quality plots and then the classification of archaea
or the classification of bacteria
```{r}
quality_MAGs=read.delim( 
"drep_output_directory_incomplete/checked_genomes/quality_report.tsv"
)
quality_MAGs$user_genome=quality_MAGs$Name
bacterial_taxonomy=read.delim(
"drep_output_directory_incomplete/classified_MAGs_MS5/classify/gtdbtk.bac120.summary.tsv"
)

archeal_taxonomy=read.delim(
"drep_output_directory_incomplete/classified_MAGs_MS5/classify/gtdbtk.ar53.summary.tsv",
)
both_taxon= rbind(bacterial_taxonomy, archeal_taxonomy) %>%full_join(
quality_MAGs  
)
plot1=ggplot(both_taxon, 
aes(fill= (Completeness) , y=classification, x=Contamination))+
  geom_point(size=6, shape=21)+scale_fill_distiller(
    palette = "RdBu", direction = 1, name="Completeness (%)"
  )+theme_bw()+ylab("")+
  xlab("Contamination (%)")

```

***Plot the results***
```{r}
png(filename = "status_derp_mags1.png", width = 1000)
print(plot1)
dev.off()
list1= list.files(pattern = "diammond_.txt", path =  "drep_output_directory_incomplete/dereplicated_genomes/",recursive =T)

```
***Merge the annotations of PlaBASE***
```{r}
path="drep_output_directory_incomplete/dereplicated_genomes/"
list2=list()
all_genomes=function(x) {
path_to_genome <- paste0(path, "/",list1[x])
file1= read.delim(path_to_genome, header = F)%>%mutate(G1=paste0(unique(list1[x])))
return(file1)
}
results  =lapply(1:length(list1), function(x)  all_genomes(x))
results2=do.call("rbind",results)%>%separate(., col=G1, into=c("Bin","PR"),
                                             sep="_out/")

results2$Id = gsub("_.*","", results2$V2)
beneficialtraits1=read.csv("diamond_the/plabase_map_info.csv",
                           header = T)

beneficialtraits2=full_join(results2, beneficialtraits1)%>%
na.omit()  

beneficialtraits2$user_genome=gsub("_out","", beneficialtraits2$Bin)
beneficialtraits3=full_join(beneficialtraits2,both_taxon )

write_csv(beneficialtraits3, 
"drep_output_directory_incomplete/dereplicated_genomes/MAGBMtraits.csv")
beneficialtraits4=dcast(beneficialtraits3,user_genome+classification~.,
value.var = "user_genome", length )
```

