---
title: "BernburgCTMP 3 YEARS"
author: "Ioannis D. Kampouris"
date: "2023-06-23"
output: pdf_document
---

I started with a script for using fastqc for quality control.
Quality control is now running.
Quality is exceptional. All samples have average above 250. 
Maybe I will do a trimming to less than 30 bases.
15.02.2023
After using grep I found that the raw reads contain barcode and primers. 
So I would recommend a 10 base trimming on 5` and a 10 bases trimming on 3`
I transferred everything into the R server, 
which runs on UBUNTU as well via R terminal.
Tested a bit on whether fastqc and multiQC work.

#############################################################
Code in Bash Shell
#############################################################
#!/bin/bash
#Author: Ioannis Kampouris
#Purpose: Use fastqc and multiqc in the R Server (UBUNTU OS)
#Dependencies a path for fastqc and 
# Attention: To download multiQC you need to have a git clone with pip3 install .

export PATH=$PATH:"/home/ioannis.kampouris/fastqc/FastQC" 
#Change accordingly to your path 
export PATH=$PATH:"/home/ioannis.kampouris/fastqc/MultiQC" 
#Change accordingly to your path 

find -name  "*raw_1*"|sed 's/^..//'  > sample_list.txt 
#Change accordingly to your path 

mkdir QC
filename='sample_list.txt'
echo Start
while read i;
   do 
   SAMPLE=$(echo ${i} | sed "s/raw_1\.fq\.gz//")
      echo "$SAMPLE"
   fastqc  ${SAMPLE}raw_2.fq.gz -o QC #You can remove the 
   #"RAW" if you want your files with NOVOGENE cleaning
    fastqc  ${SAMPLE}raw_1.fq.gz -o QC
     
done <"$filename"
 
 mkdir QC/FW
 mkdir QC/RE

mv QC/*raw_1* QC/FW
mv QC/*raw_2* QC/RE
cd QC/FW
python3 -m multiqc  .
cd ..
cd ..
cd QC/RE
python3 -m multiqc  .

After I run the fastqc/multiqc I plotted the results of average quality per read
##########################################################################
```{r, fig.height=10, fig.width=10}
library(readr)
library(tidyverse)
library(ggplot2)

set.seed(12122022)
```

#Start dada2 for the 2020 samples
```{r, fig.height=10, fig.width=10}
library(dada2)
source("mod_loess_function.R")

path_2020 <-paste0( "2020/trimmed_cutadapt")
# CHANGE ME to the directory containing your demultiplexed fastq files
filtpathF_2020 <- file.path(path_2020, "FW_filtered") 
# Filtered forward files go into the pathF/filtered/ subdirectory
filtpathR_2020 <- file.path(path_2020, "RE_filtered") 
fastqFs_2020 <- sort(list.files(path_2020, pattern="_1.fq"))
fastqRs_2020 <- sort(list.files(path_2020, pattern="_2.fq"))
# Filtering
out_2020<-filterAndTrim(fwd=file.path(path_2020, fastqFs_2020), filt=file.path(filtpathF_2020, fastqFs_2020),
              rev=file.path(path_2020, fastqRs_2020), 
              filt.rev=file.path(filtpathR_2020, fastqRs_2020),
               maxEE=c(2,2), rm.phix=TRUE, matchIDs = T,
            compress=TRUE, verbose=TRUE, multithread=TRUE)

```

Lowest output ~75.9%
```{r, fig.height=10, fig.width=10}
filtFs_2020 <- list.files(filtpathF_2020, pattern="fq", full.names = TRUE)
filtRs_2020 <- list.files(filtpathR_2020, pattern="fq", full.names = TRUE)
sample.names_2020 <- sapply(strsplit(basename(filtFs_2020), "_"), 
                            `[`, 1) # Assumes filename = samplename_XXX.fastq.gz
sample.namesR_2020 <- sapply(strsplit(basename(filtRs_2020), "_"), `[`, 1) 
# Assumes filename = samplename_XXX.fastq.gz
if(!identical(sample.names_2020, sample.namesR_2020)) stop("Forward and reverse files do not match.")else{"Proceed with Safety"}
names(filtFs_2020) <- sample.names_2020
names(filtRs_2020) <- sample.names_2020


errF_2020 <- learnErrors(filtFs_2020, nbases=1e8,
errorEstimationFunction = loessErrfun_mod1,multithread=TRUE)
errR_2020 <- learnErrors(filtRs_2020, nbases=1e8,
errorEstimationFunction = loessErrfun_mod1,multithread=TRUE)
plotErrors(errF_2020, nominalQ=TRUE)
plotErrors(errR_2020, nominalQ=TRUE)
```

```{r}
#Denoise Reads
ddF_2020 <- dada(filtFs_2020, err=errF_2020, multithread=40)
ddR_2020 <- dada(filtRs_2020, err=errR_2020, multithread=40)
```
```{r}
mergers_2020 <- mergePairs(ddF_2020, filtFs_2020, 
ddR_2020, filtRs_2020, minOverlap=7)
```
```{r}
# Construct sequence table and remove chimeras
seqtab_2020 <- makeSequenceTable(mergers_2020)
```
#Start dada2 for the 2021 samples
```{r, fig.height=10, fig.width=10}
path_2021 <-paste0( "2021/trimmed_cutadapt")
# CHANGE ME to the directory containing your demultiplexed fastq files
filtpathF_2021 <- file.path(path_2021, "FW_filtered") 
# Filtered forward files go into the pathF/filtered/ subdirectory
filtpathR_2021 <- file.path(path_2021, "RE_filtered") 
fastqFs_2021 <- sort(list.files(path_2021, pattern="_1.fq"))
fastqRs_2021 <- sort(list.files(path_2021, pattern="_2.fq"))
# Filtering
out_2021<-filterAndTrim(fwd=file.path(path_2021, fastqFs_2021), filt=file.path(filtpathF_2021, fastqFs_2021),
              rev=file.path(path_2021, fastqRs_2021),
              filt.rev=file.path(filtpathR_2021, fastqRs_2021),
               maxEE=c(2,2), rm.phix=TRUE, matchIDs = T,
              compress=TRUE, verbose=TRUE, multithread=10)

```

Lowest output ~75.9%
```{r, fig.height=10, fig.width=10}
filtFs_2021 <- list.files(filtpathF_2021, pattern="fq", full.names = TRUE)
filtRs_2021 <- list.files(filtpathR_2021, pattern="fq", full.names = TRUE)
sample.names_2021 <- sapply(strsplit(basename(filtFs_2021), "_"), `[`, 1) # Assumes filename = samplename_XXX.fastq.gz
sample.namesR_2021 <- sapply(strsplit(basename(filtRs_2021), "_"), `[`, 1) # Assumes filename = samplename_XXX.fastq.gz
if(!identical(sample.names_2021, sample.namesR_2021)) stop("Forward and reverse files do not match.")else{"Proceed with Safety"}
names(filtFs_2021) <- sample.names_2021
names(filtRs_2021) <- sample.names_2021


errF_2021 <- learnErrors(filtFs_2021,errorEstimationFunction = loessErrfun_mod1,
                         nbases=1e8, multithread=10)
errR_2021 <- learnErrors(filtRs_2021,
                         errorEstimationFunction = loessErrfun_mod1,
                         nbases=1e8, multithread=10)
plotErrors(errF_2021, nominalQ=TRUE)
plotErrors(errR_2021, nominalQ=TRUE)
```

```{r}
#Denoise Reads
ddF_2021 <- dada(filtFs_2021, err=errF_2021, multithread=10)
ddR_2021 <- dada(filtRs_2021, err=errR_2021, multithread=10)
```
```{r}
mergers_2021 <- mergePairs(ddF_2021, filtFs_2021, 
ddR_2021, filtRs_2021, minOverlap=7)
```
```{r}
# Construct sequence table and remove chimeras
seqtab_2021 <- makeSequenceTable(mergers_2021)
```

#Start dada2 for the 2022 samples
```{r, fig.height=10, fig.width=10}
path_2022 <-paste0( "2022/trimmed_cutadapt")
# CHANGE ME to the directory containing your demultiplexed fastq files
filtpathF_2022 <- file.path(path_2022, "FW_filtered") 
# Filtered forward files go into the pathF/filtered/ subdirectory
filtpathR_2022 <- file.path(path_2022, "RE_filtered") 
fastqFs_2022 <- sort(list.files(path_2022, pattern="_1.fq"))
fastqRs_2022 <- sort(list.files(path_2022, pattern="_2.fq"))
# Filtering
out_2022<-filterAndTrim(fwd=file.path(path_2022, fastqFs_2022), filt=file.path(filtpathF_2022, fastqFs_2022),
              rev=file.path(path_2022, fastqRs_2022), filt.rev=file.path(filtpathR_2022, fastqRs_2022),
               maxEE=c(2,2), rm.phix=TRUE, matchIDs = T,
              compress=TRUE, verbose=TRUE, multithread=10)

```

```{r, fig.height=10, fig.width=10}
filtFs_2022 <- list.files(filtpathF_2022, pattern="fq", full.names = TRUE)
filtRs_2022 <- list.files(filtpathR_2022, pattern="fq", full.names = TRUE)
sample.names_2022 <- sapply(strsplit(basename(filtFs_2022), "_"),
                            `[`, 1) # Assumes filename = samplename_XXX.fastq.gz
sample.namesR_2022 <- sapply(strsplit(basename(filtRs_2022), "_"),
                             `[`, 1) # Assumes filename = samplename_XXX.fastq.gz
if(!identical(sample.names_2022, sample.namesR_2022))
  stop("Forward and reverse files do not match.")else{"Proceed with Safety"}
names(filtFs_2022) <- sample.names_2022
names(filtRs_2022) <- sample.names_2022


errF_2022 <- learnErrors(filtFs_2022,errorEstimationFunction = loessErrfun_mod1,
                         nbases=1e8, multithread=10)
errR_2022 <- learnErrors(filtRs_2022, nbases=1e8,
errorEstimationFunction = loessErrfun_mod1,multithread=10)
plotErrors(errF_2022, nominalQ=TRUE)
plotErrors(errR_2022, nominalQ=TRUE)
```

```{r}
#Denoise Reads
ddF_2022 <- dada(filtFs_2022, err=errF_2022, multithread=10)
ddR_2022 <- dada(filtRs_2022, err=errR_2022, multithread=10)
```
```{r}
mergers_2022 <- mergePairs(ddF_2022, filtFs_2022, 
ddR_2022, filtRs_2022, minOverlap=7)
```

```{r}
# Construct sequence table and remove chimeras
seqtab_2022 <- makeSequenceTable(mergers_2022)
```


```{r,ig.height=10, fig.width=10}
st.three_years <- mergeSequenceTables(seqtab_2020, seqtab_2021,seqtab_2022)
table(nchar(getSequences(st.three_years)))
distribution_sequences=data.frame(D=(nchar(getSequences(st.three_years))))
distribution_sequences_plot=ggplot(distribution_sequences, aes(x=D))+geom_bar(stat = "count")
ggplot(distribution_sequences, aes(x=D))+geom_density(stat = "count")
png( "dada2_out/three_years_read_size_distribution.png", 
    width = 1000, height = 1000)
print(distribution_sequences_plot)
dev.off()
write.csv(distribution_sequences, 
          "dada2_out/three_years_read_size_distribution.csv")
#402 until 432
seqtab2 <- st.three_years[,nchar(colnames(st.three_years)) %in% 386:444]
seqtab3 <- removeBimeraDenovo(seqtab2, method="consensus", multithread=10)
```



```{r}
ASVtable=as.data.frame(t(seqtab3))%>%rownames_to_column(var="Sequence")

taxa <- assignTaxonomy(seqtab3, 
"~/silva_naive/silva_nr99_v138.1_train_set.fa.gz", minBoot = 80, outputBootstraps = T,  multithread=10)

complete_ASV_table= full_join(as.data.frame(taxa)%>%rownames_to_column(var = "Sequence"), ASVtable, by="Sequence")
complete_ASV_table$ASV=paste0("ASV", rownames(complete_ASV_table))


complete_ASV_table = complete_ASV_table%>%
  filter(tax.Kingdom!="NA"&tax.Order!="Chloroplast"&tax.Family!="Mitochondria")
ASV_counts=as.data.frame( rowSums(complete_ASV_table[,14:ncol(complete_ASV_table)]%>%column_to_rownames(var="ASV")))

complete_ASV_table2 = complete_ASV_table%>%
  filter(tax.Kingdom=="Bacteria")
complete_ASV_table_cleaned2=complete_ASV_table2


getN <- function(x) sum(getUniques(x))
track_all_three <- cbind(rbind(out_2020,out_2021,out_2022), 
 rbind(sapply(ddF_2020, getN)%>%as.data.frame() ,sapply(ddF_2021, getN)%>%as.data.frame(),sapply(ddF_2022, getN)%>%as.data.frame()), 
rbind( sapply(ddR_2020, getN)%>%as.data.frame(), sapply(ddR_2021, getN)%>%as.data.frame(), sapply(ddR_2022, getN)%>%as.data.frame()), rbind(sapply(mergers_2020, getN)%>%as.data.frame(), sapply(mergers_2021, getN)%>%as.data.frame(),sapply(mergers_2022, getN)%>%as.data.frame()), rowSums(seqtab2),rowSums(seqtab3), colSums(complete_ASV_table[,14:(ncol(complete_ASV_table)-1)]), colSums(complete_ASV_table2[,14:(ncol(complete_ASV_table)-1)]))

  
colnames(track_all_three)<-c("In", "Out","Denoised FW","Denoised RE"
                             , "Merged","Sizeout", "Chimera Filtered",
                             "Final_Clean_Of_Rare_Seq_and_Plastids", "Archea_out")
write.csv(complete_ASV_table_cleaned2,"ASV_Table_LTE_three_years.csv")

write.csv(track_all_three, file = "track_sequence_number.csv")
writeFasta<-function(data, filename){
  fastaLines = c()
  for (rowNum in 1:nrow(data)){
    fastaLines = c(fastaLines, 
as.character(paste(">", data[rowNum,"ASV"], sep = "")))
fastaLines = c(fastaLines,as.character(data[rowNum,"Sequence"]))
  }
  
  fileConn<-file(filename)
  writeLines(fastaLines, fileConn)
  close(fileConn)
}

writeFasta(complete_ASV_table_cleaned2%>%select(ASV, Sequence), "ASVs_LTE_three_years.fa")
``` 
```{r}
```