---
title: "MAG"
author: "Theresa Kuhl-Nagel"
date: "2025-03-22"
output: html_document
---

```{r}
library(ggplot2)
library(patchwork)
library(rcompanion)
library(agricolae)
MAG2 <- read.table("~/Data/Mapping_matrix_all2_reduced.txt", header=T, sep="\t", dec=".") 
head(MAG2)

new_names <- read.table("~/Data/New_names.txt", header=T, sep="\t", dec=".") 
```

Calculate relative abundance based on bacterial reads with hit in MAGs
```{r}
#separate metadata
colnames(MAG2)
meta_data <- MAG2[, c("Year", "Treatment", "Replicate")]
rest_data <- MAG2[, !colnames(MAG2) %in% c("Year", "Treatment", "Replicate")]
# View the results
print(meta_data)
print(rest_data)

#transpose rest_data
trest_data <- t(rest_data)
trest_data

#relative abundance
rowsum_MAG <- rowSums(trest_data)
rowsum_MAG
MAG_rela <- trest_data/rowsum_MAG  
View(MAG_rela) 

#transpose back
tMAG_rela <- t(MAG_rela)
tMAG_rela

#add meta_data again
MAG_relabu_final <- cbind(tMAG_rela, meta_data)
View(MAG_relabu_final)
MAG_relabu_final
```
Updated with 3 MAGs out of quality criteria removed 
```{r}
#reshape in long format
# Specify genome column names
genome_columns <- c("S4SemiBin2.fa.gzround4pilon",
"S4SemiBin19.fa.gzround4pilon",	"S4SemiBin18.fa.gzround4pilon",	"S4SemiBin1.fa.gzround4pilon",	"S4SemiBin0.fa.gzround4pilon",	"S3SemiBin0.fa.gzround4pilon",	"S1SemiBin26.fa.gzround4pilon",	"S1SemiBin16.fa.gzround4pilon",	"S1SemiBin12.fa.gzround4pilon",	"MS5ResultsCTExtC2022bin.016",	"MS5ResultsCTExtC2020bin.002",	"MS5Resultsall2bin.027",	"MS5Resultsall2bin.012",	"Genome.15")

# Select genome columns
genome_data <- MAG_relabu_final[, genome_columns]

library(tidyr)
# Reshape genome columns into long format
long_data <- MAG_relabu_final %>%
  pivot_longer(
    cols = all_of(genome_columns), # Select genome columns (columns 1–17)
    names_to = "Genome",          # Name for the new column that identifies genomes
    values_to = "Mapped_Reads"    # Name for the new column that contains mapped reads
  )

View(long_data)
```

```{r}
#calculate standard deviation
library(dplyr)

long_data <- long_data %>%
  group_by(Year, Genome, Treatment) %>%
  mutate(
    mean_reads = mean(Mapped_Reads),
    sd_value = sd(Mapped_Reads)
  ) %>%
  ungroup()
```


```{r}
# new names

long_data <- long_data %>%
  left_join(new_names, by = c("Genome"="Old.name"))
```

```{r}
custom_colors <- c( "#ff9900","#146eb4")

#change order to start with MAG01
desired_order <- sort(unique(long_data$New.name))
```

```{r}
#adapted labels
myLabels2 <- c(
  expression(paste("MAG14(", italic("Binatia"),")")),
  expression(paste("MAG13(", italic("Nitrososphaeraceae"),")")),
  expression(paste("MAG12(", italic("Nitrososphaeraceae"),")" )),
  expression(paste("MAG11(", italic("Nitrospira"),")" )),
  expression(paste("MAG10(", italic("Bacillus atrophaeus"),")" )),
  expression(paste("MAG09(", italic("Lysobacter antibioticus"),")" )),
  expression(paste("MAG08(", italic("Nitrososphaeraceae"),")")),
  expression(paste("MAG07(", italic("Nitrososphaera"),")")),
  expression(paste("MAG06(", italic("Nitrososphaeraceae"),")")),
  expression(paste("MAG05(", italic("Binatia"),")")),
  expression(paste("MAG04(", italic("Rariglobus"),")")),
  expression(paste("MAG03(", italic("Edaphocola"),")" )),
  expression(paste("MAG02(",italic("Nitrososphaeraceae"),")")),
  expression(paste("MAG01(",italic("Binatia"),")",))
)  

```

```{r}
plot_list <- lapply(sort(unique(long_data$Year)), function(current_year) {
  is_first_year <- current_year == min(long_data$Year)
  
  ggplot(subset(long_data, Year == current_year),
         aes(y = New.name, x = mean_reads, fill = Treatment)) +
    scale_fill_manual(values = custom_colors) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.6) +
    geom_errorbar(
      aes(xmin = mean_reads - sd_value, 
          xmax = mean_reads + sd_value),
      position = position_dodge(width = 0.7),
      width = 0.2,
      color = "gray30",
      linewidth = 0.3
    ) +
    scale_x_continuous(limits = c(0, 0.08), expand = expansion(mult = c(0, 0.1))) +
    scale_y_discrete(limits = rev(desired_order), labels = myLabels2)+
    labs(title = current_year,
         #x = if(is_first_year) "Mapped Reads" else "",
         #y = if(is_first_year) "" else "") +
         x = "Relative abundance (%)" ,
         y = if(is_first_year) "" else "") +
    theme_minimal(base_size = 10) +
    theme(
      axis.text.y = if(is_first_year) element_text(size = 10) else element_blank(),
      axis.title.x = element_text(size = 10),
      axis.text.x = element_text(size = 10),
      panel.grid.major.y = element_blank(),
      plot.margin = margin(5, 5, 5, if(is_first_year) 5 else -10)
    )
})

# Combine plots
combined <- wrap_plots(plot_list, nrow = 1, guides = "collect") & 
  theme(legend.position = "bottom", legend.title = element_blank(), legend.text = element_text(size = 10))

combined
```

#2-way-ANOVA for significant differences
```{r}
#Load data in a different format
MAG2 <- read.table("~/Data/Mapping_matrix_all2_reduced.txt", header=T, sep="\t", dec=".") 
head(MAG2)
str(MAG2)

#Calculate relative abundance based on bacterial reads with hit in MAGs
#separate metadata
colnames(MAG2)
meta_data <- MAG2[, c("Year", "Treatment", "Replicate")]
rest_data <- MAG2[, !colnames(MAG2) %in% c("Year", "Treatment", "Replicate")]
# View the results
print(meta_data)
print(rest_data)

#transpose rest_data
trest_data <- t(rest_data)
trest_data

#relative abundance
rowsum_MAG <- rowSums(trest_data)
rowsum_MAG
MAG_rela <- trest_data/rowsum_MAG  
View(MAG_rela) 

#transpose back
tMAG_rela <- t(MAG_rela)
tMAG_rela

#add meta_data again
MAG_relabu_final <- cbind(tMAG_rela, meta_data)
View(MAG_relabu_final)
MAG_relabu_final

#as factor
MAG_relabu_final$Year <- as.factor(MAG_relabu_final$Year)
MAG_relabu_final$Treatment <- as.factor(MAG_relabu_final$Treatment)
MAG_relabu_final$Replicate <- as.factor(MAG_relabu_final$Replicate)

```

#S4SemiBin2.fa.gzround4pilon
```{r}
data_aov <- aov(S4SemiBin2.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
Analysis of Variance Table

Response: S4SemiBin2.fa.gzround4pilon
               Df     Sum Sq    Mean Sq F value  Pr(>F)   
Treatment       1 1.6435e+10 1.6435e+10  8.4314 0.00947 **
Year            2 6.1587e+09 3.0793e+09  1.5797 0.23330   
Treatment:Year  2 1.1610e+10 5.8052e+09  2.9781 0.07633 . 
Residuals      18 3.5087e+10 1.9493e+09   

BMc:2020	483844.8	a		
BMc:2022	424901.0	ab		
control:2021	420877.2	ab		
BMc:2021	415595.7	ab		
control:2020	380028.2	b		
control:2022	365525.5	b

#S4SemiBin19.fa.gzround4pilon
```{r}
data_aov <- aov(S4SemiBin19.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # significant

# data transformation
Index =
  transformTukey(MAG_relabu_final$S4SemiBin19.fa.gzround4pilon,
                 plotit=FALSE)
data_index <- cbind(MAG_relabu_final, Index)

#neue ANOVA
data_aov_index <- aov(Index ~Treatment*Year, data_index)
anova(data_aov_index)
shapiro.test(residuals(data_aov_index)) # not significant
qqnorm(residuals(data_aov_index))
leveneTest(data_aov_index) # not significant
```
not significant

#S4SemiBin18.fa.gzround4pilon
```{r}
data_aov <- aov(S4SemiBin18.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # significant
```
not significant

#S4SemiBin1.fa.gzround4pilon
```{r}
data_aov <- aov(S4SemiBin1.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
Analysis of Variance Table

Response: S4SemiBin1.fa.gzround4pilon
               Df     Sum Sq    Mean Sq F value   Pr(>F)   
Treatment       1 6.7013e+09 6701294870 10.6257 0.004351 **
Year            2 2.6915e+09 1345750648  2.1338 0.147359   
Treatment:Year  2 4.3149e+09 2157443600  3.4209 0.055057 . 
Residuals      18 1.1352e+10  630669761  

BMc:2020	286198.0	a		
BMc:2022	250224.5	ab		
control:2021	246190.6	ab		
BMc:2021	243370.0	ab		
control:2020	223116.8	b		
control:2022	209819.8	b

#S4SemiBin0.fa.gzround4pilon
```{r}
data_aov <- aov(S4SemiBin0.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
Analysis of Variance Table

Response: S4SemiBin0.fa.gzround4pilon
               Df     Sum Sq    Mean Sq F value   Pr(>F)   
Treatment       1 2200800523 2200800523 14.3570 0.001344 **
Year            2  389733832  194866916  1.2712 0.304498   
Treatment:Year  2 1029352951  514676476  3.3575 0.057651 . 
Residuals      18 2759247421  153291523    

BMc:2020	108415.25	a		
BMc:2021	91571.67	ab		
BMc:2022	85754.25	ab		
control:2021	82291.80	ab		
control:2022	74894.25	b		
control:2020	70418.50	b

#S3SemiBin0.fa.gzround4pilon
```{r}
data_aov <- aov(S3SemiBin0.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

# data transformation
Index =
  transformTukey(MAG_relabu_final$S3SemiBin0.fa.gzround4pilon,
                 plotit=FALSE)
data_index <- cbind(MAG_relabu_final, Index)

#neue ANOVA
data_aov_index <- aov(Index ~Treatment*Year, data_index)
anova(data_aov_index)
shapiro.test(residuals(data_aov_index)) # significant
qqnorm(residuals(data_aov_index))
leveneTest(data_aov_index) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov_index,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
not significant

#S1SemiBin26.fa.gzround4pilon
```{r}
data_aov <- aov(S1SemiBin26.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant
```
not significant

#S1SemiBin16.fa.gzround4pilon
```{r}
data_aov <- aov(S1SemiBin16.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant
```
not significant

#S1SemiBin12.fa.gzround4pilon
```{r}
data_aov <- aov(S1SemiBin12.fa.gzround4pilon ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # significant

# data transformation
Index =
  transformTukey(MAG_relabu_final$S1SemiBin12.fa.gzround4pilon,
                 plotit=FALSE)
data_index <- cbind(MAG_relabu_final, Index)

#neue ANOVA
data_aov_index <- aov(Index ~Treatment*Year, data_index)
anova(data_aov_index)
shapiro.test(residuals(data_aov_index)) # not significant
qqnorm(residuals(data_aov_index))
leveneTest(data_aov_index) # not significant
```
not significant

#MS5ResultsCTExtC2022bin.016
```{r}
data_aov <- aov(MS5ResultsCTExtC2022bin.016 ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

# data transformation
Index =
  transformTukey(MAG_relabu_final$MS5ResultsCTExtC2022bin.016,
                 plotit=FALSE)
data_index <- cbind(MAG_relabu_final, Index)

#neue ANOVA
data_aov_index <- aov(Index ~Treatment*Year, data_index)
anova(data_aov_index)
shapiro.test(residuals(data_aov_index)) # significant
qqnorm(residuals(data_aov_index))
leveneTest(data_aov_index) # not significant
```
not significant

#MS5ResultsCTExtC2020bin.002
```{r}
data_aov <- aov(MS5ResultsCTExtC2020bin.002 ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # significant

# data transformation
Index =
  transformTukey(MAG_relabu_final$MS5ResultsCTExtC2020bin.002,
                 plotit=FALSE)
data_index <- cbind(MAG_relabu_final, Index)

#neue ANOVA
data_aov_index <- aov(Index ~Treatment*Year, data_index)
anova(data_aov_index)
shapiro.test(residuals(data_aov_index)) # significant
qqnorm(residuals(data_aov_index))
leveneTest(data_aov_index) # not significant
```
not significant

#MS5Resultsall2bin.027
```{r}
data_aov <- aov(MS5Resultsall2bin.027 ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
Analysis of Variance Table

Response: MS5Resultsall2bin.027
               Df     Sum Sq    Mean Sq F value    Pr(>F)    
Treatment       1 6.5092e+10 6.5092e+10 15.6268 0.0009324 ***
Year            2 6.8449e+09 3.4224e+09  0.8216 0.4555426    
Treatment:Year  2 4.0708e+10 2.0354e+10  4.8864 0.0201777 *  
Residuals      18 7.4977e+10 4.1654e+09 

BMc:2020	718977.8	a		
BMc:2022	629586.0	ab		
control:2021	592546.0	ab		
BMc:2021	591978.3	ab		
control:2022	523641.0	b		
control:2020	514450.2	b

#MS5Resultsall2bin.012
```{r}
data_aov <- aov(MS5Resultsall2bin.012 ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
Analysis of Variance Table

Response: MS5Resultsall2bin.012
               Df     Sum Sq    Mean Sq F value  Pr(>F)  
Treatment       1 1.8366e+10 1.8366e+10  7.5170 0.01341 *
Year            2 1.0123e+10 5.0616e+09  2.0716 0.15498  
Treatment:Year  2 2.4521e+10 1.2261e+10  5.0180 0.01854 *
Residuals      18 4.3980e+10 2.4433e+09  

BMc:2020	494930.0	a		
control:2021	446003.0	ab		
BMc:2022	426239.2	ab		
BMc:2021	420909.3	ab		
control:2022	362764.8	b		
control:2020	361024.5	b

#Genome.15 
```{r}
data_aov <- aov(Genome.15 ~Treatment*Year, MAG_relabu_final)
anova(data_aov) 
shapiro.test(residuals(data_aov)) # not significant
qqnorm(residuals(data_aov))
leveneTest(data_aov) # not significant

#### Tukey HSD ####
HSD.Test_data_aov <- HSD.test(data_aov,c("Treatment","Year"), group=TRUE,console=TRUE, main="Posthoc: Tukey")
```
Analysis of Variance Table

Response: Genome.15
               Df     Sum Sq    Mean Sq F value  Pr(>F)  
Treatment       1 2973822612 2973822612  6.6425 0.01898 *
Year            2 1266497182  633248591  1.4145 0.26881  
Treatment:Year  2 2147582882 1073791441  2.3985 0.11928  
Residuals      18 8058488189  447693788 

BMc:2020	224353.5	a		
BMc:2022	199487.2	ab		
control:2021	199322.8	ab		
BMc:2021	196662.0	ab		
control:2020	180239.2	ab		
control:2022	173217.0	b


