---
title: "Untitled"
author: "Theresa Kuhl-Nagel"
date: "2025-08-12"
output: html_document
---
```{r}
library(ggplot2)
library(patchwork)
MAG2 <- read.table("~/Data/Mapping_matrix_all2_reduced.txt", header=T, sep="\t", dec=".") 
head(MAG2)
```

Calculate relative abundance based on bacterial reads with hit in MAGs
```{r}
#separate metadata
colnames(MAG2)
meta_data <- MAG2[, c("Year", "Treatment", "Replicate")]
rest_data <- MAG2[, !colnames(MAG2) %in% c("Year", "Treatment", "Replicate")]
# View the results
print(meta_data)
print(rest_data)

#transpose rest_data
trest_data <- t(rest_data)
trest_data

#relative abundance
rowsum_MAG <- rowSums(trest_data)
rowsum_MAG
MAG_rela <- trest_data/rowsum_MAG  
View(MAG_rela) 

#transpose back
tMAG_rela <- t(MAG_rela)
tMAG_rela

#add meta_data again
MAG_relabu_final <- cbind(tMAG_rela, meta_data)
View(MAG_relabu_final)
MAG_relabu_final
```
```{r}
custom_colors <- c( "#ff9900","#146eb4")
```


#S4SemiBin2.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S4SemiBin2.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S4SemiBin2.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S4SemiBin2.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S4SemiBin2.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#S4SemiBin18.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S4SemiBin18.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S4SemiBin18.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S4SemiBin18.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S4SemiBin18.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#S4SemiBin1.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S4SemiBin1.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S4SemiBin1.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S4SemiBin1.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S4SemiBin1.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```
#S4SemiBin0.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S4SemiBin0.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S4SemiBin0.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S4SemiBin0.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S4SemiBin0.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```
#S3SemiBin0.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S3SemiBin0.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S3SemiBin0.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S3SemiBin0.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S3SemiBin0.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#S1SemiBin26.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S1SemiBin26.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S1SemiBin26.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S1SemiBin26.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S1SemiBin26.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#S1SemiBin16.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S1SemiBin16.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S1SemiBin16.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S1SemiBin16.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S1SemiBin16.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#S1SemiBin12.fa.gzround4pilon
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(S1SemiBin12.fa.gzround4pilon, na.rm = TRUE),
    sd = sd(S1SemiBin12.fa.gzround4pilon, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("S1SemiBin12.fa.gzround4pilon")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = S1SemiBin12.fa.gzround4pilon, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#MS5ResultsCTExtC2022bin.016
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(MS5ResultsCTExtC2022bin.016, na.rm = TRUE),
    sd = sd(MS5ResultsCTExtC2022bin.016, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("MS5ResultsCTExtC2022bin.016")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = MS5ResultsCTExtC2022bin.016, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#MS5ResultsCTExtC2020bin.002
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(MS5ResultsCTExtC2020bin.002, na.rm = TRUE),
    sd = sd(MS5ResultsCTExtC2020bin.002, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("MS5ResultsCTExtC2020bin.002")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = MS5ResultsCTExtC2020bin.002, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#MS5Resultsall2bin.027
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(MS5Resultsall2bin.027, na.rm = TRUE),
    sd = sd(MS5Resultsall2bin.027, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("MS5Resultsall2bin.027")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = MS5Resultsall2bin.027, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```
#MS5Resultsall2bin.012
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(MS5Resultsall2bin.012, na.rm = TRUE),
    sd = sd(MS5Resultsall2bin.012, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("MS5Resultsall2bin.012")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = MS5Resultsall2bin.012, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```

#Genome.15
```{r}
library(dplyr)
#standard deviation
summary <- MAG_relabu_final %>%
  group_by(Year, Treatment) %>%
  summarise(
    mean = mean(Genome.15, na.rm = TRUE),
    sd = sd(Genome.15, na.rm = TRUE)
  ) %>%
  ungroup()

# Bar plot 
p1 <- ggplot(summary, aes(x = Year, y = mean, fill = Treatment)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2
  ) +
  labs(x = "Year", y = expression("Genome.15")) +
  scale_fill_manual(values = custom_colors) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.title = element_blank()
  )

# Use only p.format as label. Remove method name. Default method = Wilcoxon test
p1<-p1 + stat_compare_means(data = MAG_relabu_final, aes(x = Year, y = Genome.15, group = Treatment),label = "p.format", label.x = 2)#size = 3, label.y = 500
#p1 + stat_compare_means(label = "p.signif")
#p1 + stat_compare_means( aes(label = ..p.signif..), label.x = 1.5, label.y = 6.5)
#p1 <- p1 + stat_compare_means(method ="anova")

p1
```


















