---
title: "CT202122_statistics_ordination"
author: "Theresa Kuhl-Nagel"
date: "2024-06-24"
output: html_document
---
```{r}
library(agricolae)
library(vegan)
#install.packages("BiocManager")
library(RColorBrewer)
library(ggplot2)
library(ellipse)
library(writexl)
```

Load data
```{r}
CDB <- read.csv("~/Data/CTExt2022122_bacteria_Plabase_anno_sum.csv", row.names = 1)
dim(CDB)
head(CDB)

metadat <- read.table("~/Data/Metadata_CTExt202122.txt", header=T, sep="\t", dec=".") 
head(metadat)
```

#switch rows and columns
```{r}
tCDB <- t(CDB) # samples in rows
dim(tCDB)
head(tCDB)
```

#Unfiltered data and relative abundance; log10 transformed
Calculate relative abundance based on bacterial reads with hit in Plabase
```{r}
#samples in rows!
rowsum_CDB <- rowSums(tCDB)
CDB_rela <- tCDB/rowsum_CDB   
View(CDB_rela) 

#tranformation
CDB_rela_trafo <- decostand(CDB_rela, method = "log", logbase = 10)
View(CDB_rela_trafo)

```

#PERMANOVA on relative abundance 
```{r}
CDB_dist <-vegdist(CDB_rela_trafo, method="bray") 
perm.ado<-adonis2(formula=CDB_dist ~ Beneficials * Year, data=metadat, by="terms", method="Bray", permutation=10000) 
perm.ado 
```

#Separate years
```{r}
CDB2020 <- CDB_rela_trafo[c(1:8), ]
head(CDB2020)
dim(CDB2020)

CDB2021 <- CDB_rela_trafo[c(9:16), ]
head(CDB2021)
dim(CDB2021)

CDB2022 <- CDB_rela_trafo[c(17:24), ]
head(CDB2022)
dim(CDB2022)

meta2020 <- metadat[c(1:8), ]
meta2021<- metadat[c(9:16), ]
meta2022<- metadat[c(17:24), ]
```

#CDB2020: PERMANOVA on relative abundance 
```{r}
CDB_dist2020 <-vegdist(CDB2020, method="bray") 
perm.ado<-adonis2(formula=CDB_dist2020~Beneficials, data=meta2020, by = "terms", method="Bray", permutation=10000) 
perm.ado
```

#CDB2021: PERMANOVA on relative abundance 
```{r}
CDB_dist2021 <-vegdist(CDB2021, method="bray") 
perm.ado<-adonis2(formula=CDB_dist2021~Beneficials, data=meta2021, by = "terms", method="Bray", permutation=10000) 
perm.ado
```

#CDB2022: PERMANOVA on relative abundance 
```{r}
CDB_dist2022 <-vegdist(CDB2022, method="bray") 
perm.ado<-adonis2(formula=CDB_dist2022~Beneficials, data=meta2022, by = "terms", method="Bray", permutation=10000) 
perm.ado
```
######################## Run plots as ggplot

2020 
```{r}
dist1=vegdist(CDB2020, method = "bray")
dist1b=cmdscale(dist1, k=2, eig=T)
dist1c=scores(dist1b, display = "sites")
dist1d=cbind(dist1c[meta2020$Sample,], meta2020)
```

```{r}
p1= ggplot(dist1d,aes(x=Dim1, y=Dim2, fill=(Beneficials) ))+
  geom_point(shape=21,  size=6)+
  #scale_shape_manual(name="Sampling Year",values = c(21,22,23))+
   theme_bw()+
  theme( strip.background = element_rect( fill="white", colour="white")) +
  theme( strip.text =  element_text( colour="black", face="bold.italic", size=20)) +
  #ggtitle(paste( "2020"))+
  scale_fill_manual(name="",values=c( "#ff9900","#146eb4"))+

  theme(axis.text = element_text(size = 10, face = "bold", colour="black")) +
  theme(legend.text = element_text( size=10, face="bold", colour = "black")) + 
      theme(plot.title = element_text( size=10, face="bold", colour = "black")) + 

    theme(legend.title = element_text( size=10, face="bold", colour = "black")) + 
  
  theme(legend.position = "right",
        axis.title.y = element_text( size=10, face = "bold", colour="black")) + 
  theme(axis.title.x = element_text(size = 10, face = "bold", colour="black"))+
  xlab(paste0("MDS1[",round((100*dist1b$eig[1]/sum(dist1b$eig)),2),"%]"))+
  ylab(paste0("MDS2[",round((100*dist1b$eig[2]/sum(dist1b$eig)),2),"%]"))

p1
```

2021
```{r}
dist1=vegdist(CDB2021, method = "bray")
dist1b=cmdscale(dist1, k=2, eig=T)
dist1c=scores(dist1b, display = "sites")
dist1d=cbind(dist1c[meta2021$Sample,], meta2021)
```

```{r}
p2= ggplot(dist1d,aes(x=Dim1, y=Dim2, fill=(Beneficials) ))+
  geom_point(shape=21,  size=6)+
  #scale_shape_manual(name="Sampling Year",values = c(21,22,23))+
   theme_bw()+
  theme( strip.background = element_rect( fill="white", colour="white")) +
  theme( strip.text =  element_text( colour="black", face="bold.italic", size=25)) +
  #ggtitle(paste( "2021"))+
  scale_fill_manual(name="",values=c( "#ff9900","#146eb4"))+

  theme(axis.text = element_text(size = 10, face = "bold", colour="black")) +
  theme(legend.text = element_text( size=10, face="bold", colour = "black")) + 
      theme(plot.title = element_text( size=10, face="bold", colour = "black")) + 

    theme(legend.title = element_text( size=10, face="bold", colour = "black")) + 
  
  theme(legend.position = "right",
        axis.title.y = element_text( size=10, face = "bold", colour="black")) + 
  theme(axis.title.x = element_text(size = 10, face = "bold", colour="black"))+
  xlab(paste0("MDS1[",round((100*dist1b$eig[1]/sum(dist1b$eig)),2),"%]"))+
  ylab(paste0("MDS2[",round((100*dist1b$eig[2]/sum(dist1b$eig)),2),"%]"))

p2
```
2022
```{r}
dist1=vegdist(CDB2022, method = "bray")
dist1b=cmdscale(dist1, k=2, eig=T)
dist1c=scores(dist1b, display = "sites")
dist1d=cbind(dist1c[meta2022$Sample,], meta2022)
```

```{r}
p3= ggplot(dist1d,aes(x=Dim1, y=Dim2, fill=(Beneficials) ))+
  geom_point(shape=21,  size=6)+
  #scale_shape_manual(name="Sampling Year",values = c(21,22,23))+
   theme_bw()+
  theme( strip.background = element_rect( fill="white", colour="white")) +
  theme( strip.text =  element_text( colour="black", face="bold.italic", size=25)) +
  #ggtitle(paste( "2022"))+
  scale_fill_manual(name="",values=c( "#ff9900","#146eb4"))+

  theme(axis.text = element_text(size = 10, face = "bold", colour="black")) +
  theme(legend.text = element_text( size=10, face="bold", colour = "black")) + 
      theme(plot.title = element_text( size=10, face="bold", colour = "black")) + 

    theme(legend.title = element_text( size=10, face="bold", colour = "black")) + 
  
  theme(legend.position = "right",
        axis.title.y = element_text( size=10, face = "bold", colour="black")) + 
  theme(axis.title.x = element_text(size = 10, face = "bold", colour="black"))+
  xlab(paste0("MDS1[",round((100*dist1b$eig[1]/sum(dist1b$eig)),2),"%]"))+
  ylab(paste0("MDS2[",round((100*dist1b$eig[2]/sum(dist1b$eig)),2),"%]"))

p3
```

Combine plots
```{r, fig.width=12, fig.height=6}

combined_plot <- ggarrange(p1, p2, p3,p4,p5,p6,
          ncol = 3, nrow = 2,
          align = "v",        
          common.legend = FALSE) 

print(combined_plot)
```




